A Switching Planner for Combined Task and Observation Planning

### 2010-11-19 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
Initial state

Because PPDDL cannot model domains that feature partial observability,
we developed an extension of that language, called DTPDDL. This can
express probabilistic models of the sensing consequences of acting to
quantitatively capture unreliability in perception. That expressive
power is achieved by incorporating perceptual analogues of fluent and
predicate signatures, and of action definitions. In detail, we have
declarations of state characterising predicate and fluent symbols as
usual. In addition, we allow two other declarative blocks, for
perceptual predicates and fluents respectively. In detail, suppose
our robot is tasked with exploring {\em locations} in order to
identify the whereabouts of a {\em visual-object}. We must describe
state and perceptual facts that model the {\em true}, resp. perceived,
locations of objects. In DTPDDL, these declarations appear as:

\small{
\begin{tabtt}
(:\=functions  ;; state fluents\\
  \> (is-in ?v - visual-object) - location )\\
(:observable-functions  ;; perceptual fluents\\
  \> (o-is-in ?v - visual-object) - location )
\end{tabtt}
}

\noindent To model the sensing capabilities of the agent, we have
operator-like descriptive elements, with preconditions in both action
and state symbols, and uniformly positive effects over perceptual
predicates -- I.e, the negation normal form of a sensing effect
formula cannot contain negation. Moreover, and without a loss of
generality, we find it convenient to separate the actional
precondition from the state precondition. For example, our mobile
robot is able to look for a visual object, such as a box of
cornflakes, at a given place. We can model that deterministic action
using the following operator schema:

@@

\noindent In a slight departure from PPDDL, we suppose that
$\pp{look-for-object}$ has an effect effect on the state. That is,
execution incurs a penalty, $3$, corresponding to the {\em cost} of
performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
reserved word}, and occurs in {\em increased} and {\em decreased}
clauses of operator schemata. In that setting, {\em reward} is
interpreted as accumulated rewards and penalties.}

@@

\small{
\begin{tabtt}
(\= :sense vision \+\\
 :parameters \= (\= ?r - robot ?v - visual-object\\
 \>\>  ?l - location) \\
 :execution \> ( \> look-for-object ?r ?v ?l) \\
 :precondition (and (= (is-in ?r) ?l) ) \\
 :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
   \> \> (probabilistic 0.8 \\
   \>  \>(= (observe-is-in ?v) ?l))) \\
  \> (when (not (= (is-in ?v) ?l)) \\
   \>  \> (probabilistic 0.1 \\
   \>  \> (= (observe-is-in ?v) ?l))))) \\
\end{tabtt}
}

\noindent Here, the {\em execution} clause is a lifted description of
the sense action-precondition -- I.e,
$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{desk}) =
\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{desk})$. We
also include a redundant state precondition, $(=(\pp{is-in}~?r)?l)$,
in order to fully demonstrate the syntax. Interpreting the schema, if
action
$(\pp{look-for-object}~\pp{robot}~\pp{cornflakes}~\pp{bedroom})$ is
executed, there is a $0.8$ chance of perceiving the cornflakes if they
are in the $\pp{bedroom}$, and otherwise a $0.1$ of perceiving them.

\noindent Nodes of the tree have one of three types: (1) atomic nodes
are labelled with a state proposition, e.g., $(=
(\pp{location}~\pp{cornflakes}) \pp{kitchen})$, (2) probabilistic
nodes, have the form $(r_1, C_1;..;r_n, C_n)$, and (3) constructive
nodes, take the form of a set of probabilistic and atomic nodes. The
root node is conjunctive, and leaves are always atomic. A convenient
way to give semantics to the tree, is to treat nodes as if they were
actions. An atomic node is an action with a positive effect
corresponding to the node label, .

>>>>

Because PPDDL cannot model domains that feature partial observability,
we developed an extension of that language, called DTPDDL. This can
express probabilistic models of the sensing consequences of acting, to
quantitatively capture unreliability in perception. That expressive
power is achieved by incorporating perceptual analogues of fluent,
predicate, and action definitions. In detail, we have declarations of
state characterising predicate and fluent symbols according to PPDDL
syntax. In addition, we allow two other declarative blocks, for
perceptual predicates and fluents respectively. For example, suppose our
robot is tasked with exploring {\em locations} in order to identify
the whereabouts of a {\em visual-object}. We must describe state and
perceptual facts that model the {\em true}, resp. perceived, locations
of objects. In DTPDDL, these declarations appear as:

\small{
\begin{tabtt}
(\=:functions  ;; state fluents\\
  \> (is-in ?v - visual-object) - location )\\
(:observable-functions  ;; perceptual fluents\\
  \> (o-is-in ?v - visual-object) - location )
\end{tabtt}
}

\noindent To model the sensing capabilities of the agent, we have
operator-like descriptive elements, with preconditions in both action
and state symbols, and uniformly positive effects over perceptual
predicates -- I.e, the negation normal form of a sensing effect
formula cannot contain negation. Whereas the effects of an {\em
operator} schema are over state propositions, and describe how the
states change under application of actions, the effects of a {\em
sense} schema are perceptual, and describe the composition of an
observation following the execution of an action. Also, and without a
loss of generality, we find it convenient to separate the actional
precondition from the state precondition. For example, our robot is
able to look for a visual object, such as a box of cornflakes, at a
given place. We can model that deterministic action using the
following operator schema:

@@

\noindent In a slight departure from PPDDL, we suppose that
$\pp{look-for-object}$ has an effect on the state. That is, its
execution incurs a penalty, $3$, corresponding to the {\em cost} of
performing the visual search.\footnote{In PPDDL {\em reward} is a {\em
reserved word}, and occurs in {\em increased} and {\em decreased}
clauses of operator schemata. In that setting, {\em reward} is
interpreted as accumulated (resp. instantaneous) rewards and
penalties.}

@@

\small{
\begin{tabtt}
(\= :sense vision \+\\
 :parameters \= (\= ?r - robot ?v - visual-object\\
 \>\>  ?l - location) \\
 :execution \> ( \> look-for-object ?r ?v ?l) \\
 :precondition (and (= (is-in ?r) ?l) ) \\
 :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
   \> \> (probabilistic 0.8 \\
   \>  \>(= (o-is-in ?v) ?l))) \\
  \> (when (not (= (is-in ?v) ?l)) \\
   \>  \> (probabilistic 0.1 \\
   \>  \> (= (o-is-in ?v) ?l))))) \\
\end{tabtt}
}

\noindent Here, the {\em execution} clause is a lifted description of
the sense action-precondition -- I.e,
$(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office}) =
\poss_\stochActions(\pp{vision}~\pp{R2D2}~\pp{box}~\pp{office})$. Above,
we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
order to fully demonstrate the syntax. Interpreting the above schema,
if action $(\pp{look-for-object}~\pp{R2D2}~\pp{box}~\pp{office})$ is
executed, there is a $0.8$ chance of perceiving the box if it is in
the $\pp{office}$, and otherwise a $0.1$ of perceiving it.

The syntax and semantics for describing initial state distributions in
DTPDDL is taken verbatim from PPDDL. It is useful to present that
factored tree-like structure here, in order to aid our discussion of
sequential planning in POMDP determinisations. The distribution is
expressed in a tree-like structure of terms. Each term is either: (1)
atomic, e.g., $(=(\pp{location}~\pp{box})~\pp{office})$, (2)
probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$, or
(3) a conjunct over probabilistic and atomic terms. The root term is
conjunctive, and the leaves are atomic. For example, the starting
distribution for $\pp{R2D2}$ might be described thusly:

\small{
\begin{tabtt}
(\=:init (= (is-in R2D2) kitchen) \+ \\
       (probabilistic \=.8 (= (is-in box) office)  \\
		      \>.2 (= (is-in box) kitchen)) \\
       (probabilistic .3 (= (is-in cup) office)  \\
		      \>.7 (= (is-in cup) kitchen))) \\
\end{tabtt}
}

\noindent The interpretation of such an expression can be given
according to a recursive descent.
